---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hurricanesotr

<!-- badges: start -->
<!-- badges: end -->

The goal of hurricanesotr is to provide tools to manipulate  hurricanes data. If given a database of recorded hurricanes, the hurricanesotr package helps produce a single 
observation time dataset for any single storm in the database. It is done by supplying to 
sotdfss function the following arguments: name, year, month, day and hour. 
The default database for storms is ext_tracks. A data downloaded from 
 \url{http://rammb.cira.colostate.edu/research/tropical_cyclones/tc_extended_best_track_dataset/} and cleaned by a code provided in data-raw. 
 wborder and wborders are the main functions called by StatHurricane to transform the data.
 stat_hurricane and geom_hurricane help plot the wind radii for single observation time dataset as polygons

## Installation

You can install the development version of hurricanesotr like so:

``` r
install.packages("devtools")
library(devtools)
devtools::install_github("hugo7gpad/hurricanesotr")
```

## Example

This is a basic example which shows you how to get the single observation time dataset
for the storm KATRINA on August 29, 2005 at 12pm

```{r example}
library(hurricanesotr)
data("ext_tracks")
katrina <- sotdfss("KATRINA", 2005, "08", "29", "12")
katrina
```

Similarly, we can get the single observation time dataset for hurricane IKE. 
But first, let's get some information about its landfall. The following code gives
a simplified data for IKE 

```{r simplData}
library(dplyr)
ike_simplData <- dplyr::filter(ext_tracks, storm_name == 'IKE',
                               year == 2008) %>%
  dplyr::select(month, day, hour, longitude, latitude,
                max_wind,eye_diameter, distance_to_land)
ike_simplData
```

Using the variable distance_to_land, we notice that the storm was over the
United States for month = 09, day = 13, hour = 12. The nearest point it was
recorded before is for month = 09, day = 13, hour = 06. We are going to use
this nearest point as the landfall point of the storm IKE. 
A google map search of the longitude( - 94.6) and latitude(29.1) obtained indicates that the 
landfall occurred near Galveston, TEXAS.
The single observation time dataset at that moment will be:

```{r sotd_ike}
ike <- sotdfss("IKE", 2008, "09", "13", "06")
ike
```

You can use geom_hurricane to plot wind radii for hurricane IKE

```{r plotting}
ggplot2::ggplot(data = ike) +
  geom_hurricane(ggplot2::aes(x = longitude, y = latitude, r_ne = ne, r_nw = nw, r_sw = sw,
                              r_se = se, fill = wind_speed, colour = wind_speed),
                 scale_radii = 1, np = 250) +
  ggplot2::scale_colour_manual(name = "Wind speed (kts)", values = c("red", "orange", 'yellow')) +
  ggplot2::scale_fill_manual(name = "Wind speed (kts)", values = c("red", "orange", 'yellow'))

```


You can also save the plot of wind radii for IKE:

```{r ike_wind_radii, echo = FALSE}
# Save IKE wind radii as pdf
pdf("ike_wind_radii")
ggplot2::ggplot(data = ike) +
  geom_hurricane(ggplot2::aes(x = longitude, y = latitude, r_ne = ne, r_nw = nw, r_sw = sw,
                              r_se = se, fill = wind_speed, colour = wind_speed),
                 scale_radii = 1, np = 250) +
  ggplot2::scale_colour_manual(name = "Wind speed (kts)", values = c("red", "orange", 'yellow')) +
  ggplot2::scale_fill_manual(name = "Wind speed (kts)", values = c("red", "orange", 'yellow'))

dev.off()
```

You can also add that plot to a base map 

```{r basemap_plot}
library(ggmap)
map_data <- get_map("Galveston", zoom = 6, maptype = "toner-background")
base_map <- ggmap(map_data, extent = "device")

base_map +
  geom_hurricane(data = ike, ggplot2::aes(x = longitude, y = latitude, r_ne = ne, r_nw = nw, r_sw = sw,
                                              r_se = se, fill = wind_speed, colour = wind_speed),
                 scale_radii = 1, np = 250) +
  ggplot2::scale_colour_manual(name = "Wind speed (kts)", values = c("red", "orange", 'yellow')) +
  ggplot2::scale_fill_manual(name = "Wind speed (kts)", values = c("red", "orange", 'yellow'))
```


Now, we can you use the scale_radii parameter to plot the wind radii only at 50%

```{r scaling}

# scale_radii = 0.5
 base_map +
  geom_hurricane(data = ike, ggplot2::aes(x = longitude, y = latitude, r_ne = ne, r_nw = nw, 
                                          r_sw = sw, r_se = se, fill = wind_speed,
                                          colour = wind_speed), scale_radii = 0.5, np = 250) +
  ggplot2::scale_colour_manual(name = "Wind speed (kts)", values = c("red", "orange",'yellow')) +
  ggplot2::scale_fill_manual(name = "Wind speed (kts)", values = c("red", "orange", 'yellow'))
```

Let's compare this last plot to the plot with default value for scale_radii, 1.

```{r compare}
# Default: scale_radii = 1
defaultScaleRadii_1 <- base_map +
  geom_hurricane(data = ike, ggplot2::aes(x = longitude, y = latitude, r_ne = ne, r_nw = nw, r_sw = sw,
                                              r_se = se, fill = wind_speed, colour = wind_speed),
                 scale_radii = 1, np = 250) +
  ggplot2::scale_colour_manual(name = "Wind speed (kts)", values = c("red", "orange", 'yellow')) +
  ggplot2::scale_fill_manual(name = "Wind speed (kts)", values = c("red", "orange", 'yellow')) +
  labs(title = "scale_radii = 1")

# scale_radii = 0.5
scaleRadii_0.5 <- base_map +
  geom_hurricane(data = ike, ggplot2::aes(x = longitude, y = latitude, r_ne = ne, r_nw = nw, r_sw = sw,
                                              r_se = se, fill = wind_speed, colour = wind_speed),
                 scale_radii = 0.5, np = 250) +
  ggplot2::scale_colour_manual(name = "Wind speed (kts)", values = c("red", "orange", 'yellow')) +
  ggplot2::scale_fill_manual(name = "Wind speed (kts)", values = c("red", "orange", 'yellow')) +
  labs(title = "scale_radii = 0.5")

gridExtra::grid.arrange(defaultScaleRadii_1, scaleRadii_0.5, ncol = 2)

```
